---
title: "Data Wrangling I Review - Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in FAS Litter data

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## `select`

Choose some columns and not others.

```{r}
select(litters_df, group, litter_number)
select(litters_df, group, gd0_weight:gd_of_birth)
```

Get rid of one column

```{r}
select(litters_df, -litter_number)
```


Renaming a column ...

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

You can rename without selecting columns

```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```


look up `?select_helpers`, look at _starts_with_, _ends_with_, _contains_, _everything_

```{r}
select(litters_df, starts_with("gd"))
```

You can just look at the one column of interest and keep everything else to the right.
```{r}
select(litters_df, litter_number, everything())
```


New function in dplyer `relocate` does the same thing (but we usually use select).

```{r}
relocate(litters_df, litter_number)
```


## `filter`

Remove or keep rows (select reorganizes **columns**)

Filter things out, leaves just 8 obs:
```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd_of_birth == 20)
```

double equal because asking question of equality, not saying it equals 20.

```{r}
filter(litters_df, !(gd_of_birth == 20))
```
! - see where this is true and where it's not true, flip both of those.
!= means not equal to


```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

```{r}
filter(litters_df, group == "Mod8")
```

Seeing two things simultaneously:
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```


## `mutate`

How much did they gain over the course of their pregnancies?
Can create and modify variables at the same time- added wait gain and made variables lowercase.
```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```



## `arrange`


```{r}
arrange(litters_df, pups_born_alive)
```
This is ascending

In console, can put `arrange(litters_df, pups_born_alive, gd0_weight)`


## `%>%`


```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight)
```

Whew. We can nest functions inside of each other but gross.

USE THE PIPE OPERATOR INSTEAD
_shift command m_

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>%
  select(-pups_survive) %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
```






































