---
title: "Data Wrangling I Review - Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
```
We're focusing on the `readr` package. 
We put the `readxl` package in later.
Also haven.

```{r}
getwd()
```

## Read in some data

We want to use a relative path, not an absolute path.

Read in the litters dataset.

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```


## Take a look at the data

Printing in the console.
```{r}
litters_df
```

It's telling us we have 49 rows and 8 columns, it's a tibble, variable names have been converted to snakecase compared to before. 


Beginning or end of dataset:
```{r}
head(litters_df)
tail(litters_df)
```

Summary tool:
```{r}
skimr::skim(litters_df)
```


Can type `view` in the CONSOLE. It opens up another window to see full dataset.


Type `?read_csv` in console to learn more


## Options to read_csv

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
```

```{r}
litters_df = read_csv("./data/FAS_litters.csv", na = c("", "NA", 9999, "."))
```

## Other file formats

Read in Excel file.

```{r}
library(readxl)
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df
```

Can `view(mlb_df)` in console.


Read in a SAS file.

```{r}
library(haven)
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```

## Comparison with Base R

don't ever use `read.csv`
```{r}
pups_base = read.csv("./data/FAS_pups.csv")
pups_readr = read_csv("./data/FAS_pups.csv")

pups_base
pups_readr
```


# Exporting data

Export the mlb sub-table.

```{r}
write_csv(mlb_df, "./data/mlb_subtable.csv")
```










